Automatic Annotation of Tasks in Structured Code

### 2018-07-26 docs/ACM/paper.tex
Adding URL of page.

%% For double-blind review submission
\documentclass[sigplan,10pt,review]{acmart}
\settopmatter{printfolios=true}
%% For final camera-ready submission
%\documentclass[acmlarge]{acmart}\settopmatter{}

@@

\author{Divino C\'{e}sar, Guido Ara\'{u}jo}
%\authornote{with author1 note}
%\orcid{nnnn-nnnn-nnnn-nnnn}
\affiliation{
%  \position{Researcher}
%  \department{IC}
  \institution{UNICAMP, Brazil}
%  \streetaddress{Cidade Universit\'{a}ria}
%  \city{Campinas}
%  \state{S\~{a}o Paulo}
%  \postcode{13083-970}
%  \country{Brazil}
}
\email{divcesar@gmail.com}
\email{guido@ic.unicamp.br}

@@

\author{Fernando Magno Quint\~{a}o Pereira}
\authornote{This work is sponsored by CNPq, CAPEs, FAPEMIG, FAPESP, LG Electronics and
the CONTINUUM ANR Project.}          %% \authornote is optional;
%\orcid{nnnn-nnnn-nnnn-nnnn}
\affiliation{
%  \position{Professor}
%  \department{DCC}
  \institution{UFMG, Brazil}
%  \streetaddress{6627 Antonio Carlos Avenue}
%  \city{Belo Horizonte}
%  \state{Minas Gerais}
%  \postcode{31.270-213}
%  \country{Brazil}
}
\email{fernando@dcc.ufmg.br}          %% \email is recommended

@@

\maketitle

@@

% Explain rational
The analyses described in Sections~\ref{sub:sra}-\ref{sub:variance} have
been implemented at the level of the LLVM intermediate representation.
We chose to implement those techniques at that level to benefit from the support
of several data-flow analyses already in place in the LLVM infra-structure.
In particular, this decision let us reuse LLVM's scalar
evolution~\cite[p.18]{Grosser12}, from which we derive our symbolic analysis.
Additionally, we could reuse LLVM's program dependence graph, and its suite of
powerful alias analyses.
This toolbox gave us the necessary means to disambiguate pointers, albeit
conservatively, and estimate safe limits to memory regions.
However, our annotations are still inserted in C programs.
Having them in a high-level programming language has two main advantages:
they are human-readable, and thus can be debugged, and they are compatible with
compilers other than \textsf{clang}, such as \textsf{gcc}.
Thus, a major part of this work was the engineering effort to map
information from LLVM's intermediate representation back into C.

@@

\noindent
\textbf{Benchmarking:}
we have used different benchmarks in our experiments.
In Sections~\ref{sub:performance} and~\ref{sub:optimizations} we use
\textsf{BSC-Bots} (\url{https://github.com/bsc-pm/bots/tree/master/serial})~\cite{Duran09} and
\textsf{Swan} (\url{http://cuda.dcc.ufmg.br/swan/})~\cite{Moreira17}.
In Section~\ref{sub:versatility} we use the LLVM test suite, and in
Section~\ref{sub:scalability} we use random programs produced with
\textsf{CSmith}~\cite{Yang11}.
Runtimes are averages of five executions.
No benchmark has been changed manually.

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{images/TM_Runtime}
\caption{Runtime of \Taskminer{} vs size of input programs.
Each dot represents one of the 100 program generated by \textsf{CSmit}.
We have fit a degree-2 polynomial on this dataset.
We cannot control the size of programs produced by CSmith; hence, it is
difficult to push the limit above 5,000 LoC.}
\label{fig:TM_Runtime}
\end{center}
\end{figure}

@@

This paper has described a methodology to annotate programs with task
parallel pragmas, which we could demonstrate to be effective
on general programs.
This methodology does not introduce any fundamentally new static analysis or code 
optimization; in this field, we claim no contribution.
Instead, our contributions lay into the overall design that eventually
emerged from a two-years long effort to combine existing compilation techniques
towards the goal of profiting from the powerful runtime system that OpenMP's
task parallelism brings forward.
During this period, we went through many dead-ends; too many, indeed, for a
11-pages report.
Nevertheless, this experience lets us single out a few elements from the
compiler literature, such as symbolic range analysis and windmills, which we
could use to solve challenges related to automatic code annotation.
Much work is still left to be done, until we can reach a stage in which
automatic annotations can beat consistently manual interventions.
In particular, our methodology asks for more aggressive tuning strategies.
Techniques such as Trancoso's~\cite{Trancoso17}, Iwasaki's~\cite{Iwasaki16B}
or Emani's~\cite{Emani15} have already been shown to be effective in this
domain, and we want to explore them further.
More specifically, a future release of \Taskminer{} is already expected to
feature Iwasaki's cut-off~\cite{Iwasaki16B}.

>>>>

%% For double-blind review submission
\documentclass[sigplan,10pt,screen]{acmart}
%\documentclass[sigplan,10pt,review]{acmart}
\settopmatter{printfolios=true}
%% For final camera-ready submission
%\documentclass[acmlarge]{acmart}\settopmatter{}

@@

\author{Divino Soares, Guido Ara\'{u}jo}
%\authornote{with author1 note}
%\orcid{nnnn-nnnn-nnnn-nnnn}
\affiliation{
%  \position{Researcher}
%  \department{IC}
  \institution{UNICAMP, Brazil}
%  \streetaddress{Cidade Universit\'{a}ria}
%  \city{Campinas}
%  \state{S\~{a}o Paulo}
%  \postcode{13083-970}
%  \country{Brazil}
}
\email{divcesar@gmail.com}
\email{guido@ic.unicamp.br}

@@

\author{Fernando Magno Quint\~{a}o Pereira}
\authornote{This work is sponsored by CNPq, CAPEs, FAPEMIG, FAPESP, LG Electronics and
the CONTINUUM ANR Project.}          %% \authornote is optional;
%\orcid{nnnn-nnnn-nnnn-nnnn}
\affiliation{
%  \position{Professor}
%  \department{DCC}
  \institution{UFMG, Brazil \& LIRMM, France}
%  \streetaddress{6627 Antonio Carlos Avenue}
%  \city{Belo Horizonte}
%  \state{Minas Gerais}
%  \postcode{31.270-213}
%  \country{Brazil}
}
\email{fernando@dcc.ufmg.br}          %% \email is recommended

@@

\maketitle
\renewcommand{\shortauthors}{Ramos, P., Souza, G., Soares, D., Ara\'{u}jo, G., Quint\~{a}o, F.}

@@

% Explain rational
The analyses described in Sections~\ref{sub:sra}-\ref{sub:variance} have
been implemented at the level of the LLVM intermediate representation, and can be used
through an online interface\footnote{\Taskminer{} can be
used at \url{http://cuda.dcc.ufmg.br/taskminer/}}.
We chose to implement those techniques at that level to benefit from the support
of several data-flow analyses already in place in the LLVM infra-structure, such as
scalar evolution and symbolic execution.
%In particular, this decision let us reuse LLVM's scalar
%evolution~\cite[p.18]{Grosser12}, from which we derive our symbolic analysis.
%Additionally, we could reuse LLVM's program dependence graph, and its suite of
%powerful alias analyses.
This toolbox gave us the necessary means to disambiguate pointers
and estimate safe limits to memory regions.
Yet, our annotations are still inserted in C programs.
Having them in a high-level programming language has two advantages:
readability and portability across compilers.
However, this modus operandi also brings two challenges.
First, we cannot annotate partial code: to produce LLVM bitcodes, we need all the type
definitions used in the target program.
Second, we need to map information from LLVM's intermediate representation back into C.
A type inference engine for C let us deal with the first
challenge~\cite{Melo18}.
In this section we explain how we deal with the second.

@@

\noindent
\textbf{Benchmarking:}
%we have used different benchmarks in our experiments.
In Sections~\ref{sub:performance} and~\ref{sub:optimizations} we use
\textsf{BSC-Bots} (\url{https://github.com/bsc-pm/bots/tree/master/serial})~\cite{Duran09} and
\textsf{Swan} (\url{http://cuda.dcc.ufmg.br/swan/})~\cite{Moreira17}.
In Section~\ref{sub:versatility} we use the LLVM test suite, and in
Section~\ref{sub:scalability} we use random programs produced with
\textsf{CSmith}~\cite{Yang11}.
Runtimes are averages of five executions.
%No benchmark has been changed manually.

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{images/TM_Runtime}
\caption{Runtime of \Taskminer{} vs size of input programs.
%Each dot represents one of the 100 program generated by \textsf{CSmit}.
We have fit a degree-2 polynomial on this dataset.
We cannot control the size of programs produced by CSmith; hence, it is
difficult to push the limit above 5,000 LoC.}
\label{fig:TM_Runtime}
\end{center}
\end{figure}

@@

This paper has described a methodology to annotate programs with task
parallel pragmas, which we could demonstrate to be effective
on general programs.
This methodology does not introduce any fundamentally new static analysis or code 
optimization; in this field, we claim no contribution.
Instead, our contributions lay into the overall design that eventually
emerged from a two-years long effort to combine existing compilation techniques
towards the goal of profiting from the powerful runtime system that OpenMP's
task parallelism brings forward.
During this period, we went through many dead-ends; too many, indeed, for a
11-pages report.
Nevertheless, this experience lets us single out a few elements from the
compiler literature, such as symbolic range analysis and windmills, which we
could use to solve challenges related to automatic code annotation.
Much work is still left to be done, until we can reach a stage in which
automatic annotations can beat consistently manual interventions.
In particular, our methodology asks for more aggressive tuning strategies.
Techniques such as Trancoso's~\cite{Trancoso17}, Iwasaki's~\cite{Iwasaki16B}
or Emani's~\cite{Emani15} have already been shown to be effective in this
domain, and we want to explore them further.
More specifically, a future release of \Taskminer{} is already expected to
feature Iwasaki's cut-off~\cite{Iwasaki16B}. \\
\textbf{Software} available at \url{http://cuda.dcc.ufmg.br/taskminer/}
